---
title: "Prettier Paper Repl"
output: html_document
---

---
title: "Paper Replication"
header-includes:
  - \usepackage{amsmath}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(dplyr)
library(ggplot2)
```

# Make simulated spp data
I'm trying to figure out the spatstat package here.

```{r}
library(spatstat)
library(MASS)
library(devtools)
devtools::load_all("~/Dropbox/Theses/pstallworth-thesis/spptrees")
data(plantBogs)
set.seed(1234)

#set the desired window. 
my_window <- owin(xrange = c(-1, 12), yrange = c(-1, 12))

my_lattice <- cbind(c(rep(1, 10), rep(2, 10), rep(3, 10), rep(4, 10), rep(5, 10),rep(6, 10), rep(7, 10), rep(8, 10), rep(9, 10), rep(10, 10)), rep(seq(1, 10), 10))
```

#Simulation A
```{r}
#Run simulation with homogenous poisson process with intensity lambda = 0.5
simulation_a <- rpoispp(0.5, win = my_window)
```

```{r}
mypts <- drop_points(50000, xmax = 12, xmin = -1, ymax = 12, ymin = -1)

final <- NA
final <- cbind(val_sets(10, nrow(my_lattice)), my_lattice)
d_values <- matrix(rep(NA, 3 * nrow(my_lattice)), nrow = nrow(my_lattice))
for(i in 1:nrow(my_lattice)){
  d_values[i,] <- nearest_point(my_lattice[i, 1], my_lattice[i, 2], cbind(simulation_a$x, simulation_a$y))
}
final <- cbind(final, d_values)
myRegion <- mypts[!is.na(in_Area(mypts, final)),]
myArea <- nrow(myRegion)/nrow(mypts) * 13^2
X <- seq(-1, 12, by = 1)
Y <- seq(-1, 12, by = 1)
XforZ <- as.vector(sapply(Y, rep, length(Y)))
Z <- matrix(mapply(lambda_hat, XforZ, Y, MoreArgs = list(my_set = final, bw = .536, my_region = myRegion, my_Area = myArea)), nrow = 14, byrow = T)
contour(X, Y, Z)
points(simulation_a)
```


#Simulation B
```{r}
#Run simulation with homogenous poisson process with intensity lambda = 4
simulation_b <- rpoispp(4, win = my_window )
```

```{r}
mypts <- drop_points(5000, xmax = 12, xmin = -1, ymax = 12, ymin = -1)

final <- NA
final <- cbind(val_sets(10, nrow(my_lattice)), my_lattice)
d_values <- matrix(rep(NA, 3 * nrow(my_lattice)), nrow = nrow(my_lattice))
for(i in 1:nrow(my_lattice)){
  d_values[i,] <- nearest_point(my_lattice[i, 1], my_lattice[i, 2], cbind(simulation_b$x, simulation_b$y))
}
final <- cbind(final, d_values)
myRegion <- mypts[!is.na(in_Area(mypts, final)),]
myArea <- nrow(myRegion)/nrow(mypts) * 13^2
lambda_hat(4, 4, final, 10, myRegion, myArea)
Z <- matrix(mapply(lambda_hat, XforZ, Y, MoreArgs = list(my_set = final, bw = 1.571, my_region = myRegion, my_Area = myArea)), nrow = 14, byrow = T)
contour(X, Y, Z)
```

#Simulation C
The next two simulations are slight more complicated and will require more code. The first is given by $\lambda(x, y) = 100 \times \text{ the } N_2({5 \choose 5}\left( \begin{smallmatrix} 3&0.5\sqrt{6}\\ 0.5\sqrt{6}&2 \end{smallmatrix} \right))$. First I make a function, mvdnorm, that computes the density for a bivariate gaussian distribution. Then I make the simulation.
```{r}
mvdnorm <- function(x, y, mu, sigma){
  p <- c(x, y)
  (2 * pi)^{-1} * det(sigma)^{-1 / 2} * exp(-1 / 2 * (t(p-mu) %*% solve(sigma) %*% (p-mu)))
}

mvdnorm_for_rpoispp <- function(x, y, mu, sigma){
  d <- rep(NA, length(x))
  for(i in 1:length(d)){
    d[i] <- mvdnorm(x[i], y[i], mu, sigma)
  }
  return(d)
}
my_mu <- c(5, 5)
my_sigma <- matrix(c(3, 0.5*sqrt(6), 0.5*sqrt(6), 2), nrow = 2, byrow = T)
lambda_c <- function(x, y) {100 * mvdnorm_for_rpoispp(x, y, my_mu, my_sigma)}

simulation_c <-  rpoispp(lambda = lambda_c, win = my_window)

mypts <- drop_points(5000, xmax = 12, xmin = -1, ymax = 12, ymin = -1)

final <- NA
final <- cbind(val_sets(10, nrow(my_lattice)), my_lattice)
d_values <- matrix(rep(NA, 3 * nrow(my_lattice)), nrow = nrow(my_lattice))
for(i in 1:nrow(my_lattice)){
  d_values[i,] <- nearest_point(my_lattice[i, 1], my_lattice[i, 2], cbind(simulation_c$x, simulation_c$y))
}
final <- cbind(final, d_values)
myRegion <- mypts[!is.na(in_Area(mypts, final)),]
myArea <- nrow(myRegion)/nrow(mypts) * 13^2
#simulation_c_optBW <- opt_bw(npts = 1000, nvalsets = 10, d = final,  xmax, xmin, ymax, ymin, max_guess = 2, min_guess = 0.1, step_value = 0.1)
Z <- matrix(mapply(lambda_hat, XforZ, Y, MoreArgs = list(my_set = final, bw = 0.555, my_region = myRegion, my_Area = myArea)), nrow = 14, byrow = T)
contour(X, Y, Z)
points(simulation_c, col = "red")
```

```{r}
C_info <- data.frame(c(1, 1.5, 1.5, 1, 1, 1.5, 5.5, 6, 6), c(10, 10, 9.5, 5, 4.5, 4.5, 5.5, 5.5, 6), rep(0.5, 9), rep(NA, 9))
colnames(C_info) <- c("X", "Y", "lambda", "lambda_hat")

est_lambda <- function(bw, lambda, window, lattice){
  simulation <-  rpoispp(lambda = lambda_c, win = my_window)
  mypts <- drop_points(5000, xmax = 12, xmin = -1, ymax = 12, ymin = -1)
  final <- NA
  final <- cbind(val_sets(10, nrow(lattice)), lattice)
  d_values <- matrix(rep(NA, 3 * nrow(lattice)), nrow = nrow(lattice))
  for(i in 1:nrow(lattice)){
    d_values[i,] <- nearest_point(lattice[i, 1], lattice[i, 2], cbind(simulation$x, simulation$y))
  }
  final <- cbind(final, d_values)
  myRegion <- mypts[!is.na(in_Area(mypts, final)),]
  myArea <- nrow(myRegion)/nrow(mypts) * (max(lattice[,1]) - min(lattice[,1])) * (max(lattice[,2]) - min(lattice[,2]))
  final <- final[find_duplicates(final),]
}
```
#Simulation D
The final function is just a drop off function. It is inhomogenous with $\lambda(x,y) = 0.2$ for $x <6$  and $\lambda(x, y) = 4$, otherwise. 

```{r}
lambda_d <- function(x, y){
  l <- rep(.4, length(x))
  for(i in 1:length(l))
    if(x[i]<6) l[i] = 0.2
  l
}    

simulation_d <- rpoispp(lambda_d, win = my_window)

final <- NA
final <- cbind(val_sets(10, nrow(my_lattice)), my_lattice)
d_values <- matrix(rep(NA, 3 * nrow(my_lattice)), nrow = nrow(my_lattice))
for(i in 1:nrow(my_lattice)){
  d_values[i,] <- nearest_point(my_lattice[i, 1], my_lattice[i, 2], cbind(simulation_d$x, simulation_d$y))
}
final <- cbind(final, d_values)
myRegion <- mypts[!is.na(in_Area(mypts, final)),]
myArea <- nrow(myRegion)/nrow(mypts) * 13^2
Z <- matrix(mapply(lambda_hat, XforZ, Y, MoreArgs = list(my_set = final, bw = 0.691, my_region = myRegion, my_Area = myArea)), nrow = 14, byrow = T)
contour(X, Y, Z)
points(simulation_d, col = "red")
```

#Working with the bog data
```{r read-dat, message = FALSE, fig.height = 5, fig.width = 5}
X <- seq(0, 11, by = 1)
Y <- seq(0, 11, by = 1)
XforZ <- as.vector(sapply(Y, rep, length(Y)))

d <- plantBogs

d_12M <-
  d %>%
  filter(bog == "12M" & plantX > 500 & plantX < 10500 & plantY > 500 & plantX < 10500)

d_BPB <-
  d %>%
  filter(bog == "BPB")

d_LBC <-
  d %>%
  filter(bog == "LBC")

sx <- rep(1:10, 10)
sy <- rep(1:10, each = 10)
plot(sx, sy, xlab = "", ylab = "", pch = 16)
points(d_BPB[,2] / 1000, d_BPB[,3] / 1000, col = "red")
points(d_LBC[,2] / 1000, d_LBC[,3] / 1000, col = "green")
final <- NA
final <- cbind(val_sets(10, nrow(my_lattice)), my_lattice)
d_values <- matrix(rep(NA, 3 * nrow(my_lattice)), nrow = nrow(my_lattice))
for(i in 1:nrow(my_lattice)){
  d_values[i,] <- nearest_point(my_lattice[i, 1], my_lattice[i, 2], cbind(d_12M$plantX / 1000, d_12M$plantY / 1000))
}
final <- cbind(final, d_values)

mypts <- drop_points(10000, xmax = 11, xmin = 0, ymax = 11, ymin = 0)
myRegion <- mypts[!is.na(in_Area(mypts, final)),]
myArea <- nrow(myRegion)/nrow(mypts) * 11^2
final <- final[find_duplicates(final),]
Z <- matrix(mapply(lambda_hat, XforZ, Y, MoreArgs = list(my_set = final, bw = 0.67, my_region = myRegion, my_Area = myArea)), nrow = length(X), byrow = T)
contour(X, Y, Z,xlim = c(1, 10), ylim = c(1, 10), zlim = c(0, 10))
points(my_lattice, col = "black", pch = 16)
points(d_12M$plantX / 1000, d_12M$plantY / 1000, pch = 16, col = "red")
```

An issue arises when considering how they calculated the area.
```{r}
d_EBB <-
  d %>%
  filter(bog == "EBB")
sx <- rep(1:10, 10)
sy <- rep(1:10, each = 10)
plot(sx, sy, xlab = "", ylab = "", pch = 16)
points(d_EBB[,2] / 1000, d_EBB[,3] / 1000, col = "red")

final <- NA
final <- cbind(val_sets(10, nrow(my_lattice)), my_lattice)
d_values <- matrix(rep(NA, 3 * nrow(my_lattice)), nrow = nrow(my_lattice))
for(i in 1:nrow(my_lattice)){
  d_values[i,] <- nearest_point(my_lattice[i, 1], my_lattice[i, 2], cbind(d_EBB$plantX / 1000, d_EBB$plantY / 1000))
}
final <- cbind(final, d_values)

mypts <- drop_points(10000, xmax = 11, xmin = 0, ymax = 11, ymin = 0)
myRegion <- mypts[!is.na(in_Area(mypts, final)),]
myArea <- nrow(myRegion)/nrow(mypts) * 11^2
final <- final[find_duplicates(final),]
Z <- matrix(mapply(lambda_hat, XforZ, Y, MoreArgs = list(my_set = final, bw = 0.56, my_region = myRegion, my_Area = myArea)), nrow = length(X), byrow = T)
contour(X, Y, Z, levels = seq(0, 2.4, by = 0.2), xlim = c(1, 10), ylim = c(1, 10), zlim = c(0, 5))
points(my_lattice, col = "black", pch = 16)
points(d_EBB$plantX / 1000, d_EBB$plantY / 1000, pch = 16, col = "red")
```


```{r}
d_NO5 <-
  d %>%
  filter(bog == "NO5")
sx <- rep(1:10, 10)
sy <- rep(1:10, each = 10)
plot(sx, sy, xlab = "", ylab = "", pch = 16)
points(d_NO5[,2] / 1000, d_NO5[,3] / 1000, col = "red", pch = 16)

final <- NA
final <- cbind(val_sets(10, nrow(my_lattice)), my_lattice)
d_values <- matrix(rep(NA, 3 * nrow(my_lattice)), nrow = nrow(my_lattice))
for(i in 1:nrow(my_lattice)){
  d_values[i,] <- nearest_point(my_lattice[i, 1], my_lattice[i, 2], cbind(d_NO5$plantX / 1000, d_NO5$plantY / 1000))
}
final <- cbind(final, d_values)

mypts <- drop_points(10000, xmax = 11, xmin = 0, ymax = 11, ymin = 0)
myRegion <- mypts[!is.na(in_Area(mypts, final)),]
myArea <- nrow(myRegion)/nrow(mypts) * 11^2
final <- final[find_duplicates(final),]
Z <- matrix(mapply(lambda_hat, XforZ, Y, MoreArgs = list(my_set = final, bw = 0.60, my_region = myRegion, my_Area = myArea)), nrow = length(X), byrow = T)
contour(X, Y, Z, levels = seq(0, 1, by = 0.05), xlim = c(1, 10), ylim = c(1, 10), zlim = c(0, 5))
points(my_lattice, col = "black", pch = 16)
points(d_NO5$plantX / 1000, d_NO5$plantY / 1000, pch = 16, col = "red")
```
